---
title: "Code Lab 1"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(data.table)
library(dplyr)
knitr::opts_chunk$set(echo = FALSE)
```

## A simple warm-up

I have a data.frame. I want to delete several columns by name, as simply as possible.

*Remember: you can check your work by hitting the "Run Code" button at the top right of the code window.*

```{r warm_up, exercise=TRUE, exercise.lines = 10}
DF <- data.frame(
  x=1:10,
  y=10:1,
  z=rep(5,10),
  a=11:20
)
```

<div id="warm_up-hint">
    There are a gajillion ways of doing this. How many can your group come up with?
</div>

## Problem 1


### Re-arranging a dataset

I have generated some occurrence points for three simulated species, but the data are in a weird format. I have a named list with each entry corresponding to a species. Each entry contains three vectors: one with x coordinates, one with y coordinates, and one with the name of the species. I want to combine all of these data into a single data.frame with three columns: "Species", "x", and "y" in that order.

*No cheating! Do this programmatically without editing the `dfList` input! You can edit the code however you like.*

```{r problem_1, exercise=TRUE, exercise.lines = 25}
# Here is the original format of the data
dfList <- list(
  ThunnusThynnus = list(x = c(-97.5, -174.5, 156.5, -172.5, 111.5, -132.5, 142.5, 115.5, 146.5, 
                              171.5, 151.5, 162.5), 
                        y = c(86.5, 79.5, 59.5, 77.5, 77.5, 75.5, 58.5, 82.5, 84.5, 74.5, 78.5, 
                              74.5), 
                        Species = "ThunnusThynnus"), 
RanaRana = list(x = c(170.5, -160.5, 150.5, 165.5, 78.5, 74.5, -161.5, -129.5, -134.5, -164.5, 
                      166.5, -169.5),
                y = c(86.5, 76.5, 57.5, 70.5, 77.5, 78.5, 74.5, 74.5, 72.5, 80.5, 84.5, 85.5), 
                Species = "RanaRana"), 
UpupaEpops = list(x = c(179.5, 149.5, -153.5, 129.5, 158.5, 169.5, 131.5, -147.5, -140.5, 128.5, 
                        173.5, -172.5), 
                  y = c(63.5, 57.5, 72.5, 80.5, 50.5, 58.5, 76.5, 72.5, 72.5, 81.5, 58.5, 88.5), 
                  Species = "UpupaEpops"))

# Here is how I tried to get what I want.
tempList <- c()
for (j in seq_along(dfList)){
   tempList <- append(tempList, data.table::as.data.table(c(dfList[[j]][3], dfList[[j]][1], dfList[[j]][2])))
}
assign(paste(name, "RandomPoints", sep = ""), rbindlist(tempList))
```

<div id="problem_1-hint">
    Hint: We have loaded the data.table and dplyr libraries, if either of those might be helpful. It *is* possible to fix this code without either package, though.
</div>

## Problem 2

### Sorting a dataset by multiple columns
I want to sort a data.frame by multiple columns. For example, with the data.frame below I would like to sort by column z (descending) then by column b (ascending). How do I do this?

```{r problem_2, exercise=TRUE, exercise.eval=TRUE, exercise.lines = 10}
dd <- data.frame(b = factor(c("Hi", "Med", "Hi", "Low"), 
      levels = c("Low", "Med", "Hi"), ordered = TRUE),
      x = c("A", "D", "A", "C"), y = c(8, 3, 9, 9),
      z = c(1, 1, 1, 2))
dd
```

<div id="problem_2-hint">
    Hint: We have loaded the dplyr and data.table libraries, but as with problem 1, it is possible to solve this without either package (in this case, using order()).
</div>

Aaaaand if you need some additional help...[here's a dplyr review from last spring](https://globedatasci-ku.github.io/data_wrangling/datawrangling_tidyverse.html#1)

## Problem 3

### A loop inside a function

I am trying to get more familiar with function writing and ran into a problem that I've been trying to solve for 3 hours now. Here's the code:

```{r problem_3_chunk_1, eval=T, echo = T}
test <- function (x) {
    for (i in x)
            print(x[i])
}
```

When I assign a few variables:

```{r problem_3_chunk_2, eval=T, echo = T}
a <- c(0,1)
b <- c(1,2)
```

b works fine, but 'a' gets all screwed up:

```{r problem_3_chunk_3, eval=T, echo = T, hold = F}
test(b)

test(a)
```

I figured that something is wrong with the way R is handling the ones and zeros in 'a'. but when I iterate the function at the command line, there are no issues.

```{r problem_3_chunk_4, eval=T, echo = T, hold = F}
print(a[1])

print(a[2])
```

How do I fix this?

```{r problem_3, exercise=TRUE, exercise.eval=TRUE, exercise.lines = 15}
test <- function (x) {
    for (i in x)
            print(x[i])
}

a <- c(0,1)
b <- c(1,2)

test(a)

test(b)

```

<div id="problem_3-hint">
    Hint: Keep *indices* and *values* straight.
</div>